---
title: 'STA304'
author: "Neil Montgomery"
date: "2016-03-14"
output: 
  ioslides_presentation: 
    css: 'styles.css' 
    widescreen: true 
    transition: 0.001
---
\newcommand{\E}[1]{E{\left(#1\right)}}
\newcommand{\flist}[2]{\{#1_1, #1_2, \ldots, #1_#2\}}
\newcommand{\fulist}[3]{\{#1_{{#2}1}, #1_{{#2}2}, \ldots, #1_{{#2}{#3}}\}}
\renewcommand{\bar}[1]{\overline{#1}}
\newcommand{\SE}[1]{\sqrt{\hat{V}(#1)}}

# ratio, regression, difference estimation

## recap { .build }

If one has a sample $\{(y_1,x_1),\ldots,(y_n,x_n)\}$ the following may be of interest:

1. To estimate the population ratio $R = \tau_y\big/\tau_x$ using $\hat R=r=\bar y\big/\bar x$.

2. To enable estimation of $\tau_y$ when $N$ is unknown.

3. To enable improved estimation of $\tau_y$ or $\mu_y$ when $y$ and $x$ are correlated.

Formula summary:

$r$ | $\hat\tau_y$ | $\hat\mu_y$ | $\hat V(r)$
----|--------------|-------------|------------
$\bar y\big/\bar x$ | $r\tau_x$ | $r\mu_x$ | $s^2_r$

The error bounds are based on this new object:
$$s^2_r = \frac{\sum_{i=1}^n (y_i - r x_i )^2}{n-1}$$

## $r_i = y_i - r x_i$ is a "residual"

```{r, echo=FALSE, message=FALSE}
library(rio)
library(dplyr)
library(knitr)
salary <- import("Ch 6 Exercises.xls", sheet="6.10")
N_teacher <- 750
r <- sum(salary$Present)/sum(salary$Past)

```

Reconsider the teacher salary example with $r = `r r`$. First 5 lines shown:

```{r, echo=FALSE}
salary %>% 
  mutate("Predicted = r*Past" = r*Past, r_i = Present - r*Past) %>% 
  slice(1:5)
```

Think of $rx_i$ as the "predicted" value for $y_i$. Then $y_i - rx_i$ is in some sense a prediction error, or "residual". And $s^2_r$ is just the "average" of the squared residuals.

## "residuals" plotted

```{r, echo=FALSE, fig.align='center'}
library(ggplot2)
salary %>% 
  ggplot(aes(x=Past, y=Present)) + 
  geom_point() + 
  geom_abline(slope=r, intercept = 0) + 
  geom_segment(aes(x=Past, y=Present, xend = Past, yend=r*Past), color="red")
```

## example of improved estimation { .build }

We have estimated a ratio and estimated $\tau_y$ using ratio techniques because there was no other option. In this example we'll estimate $\mu_y$ using ratio techniques simply to take advantage of the information contained in the $x$ variable.

```{r, echo=FALSE}
rats <- import("Ch 6 Exercises.xls", sheet="6.6")
N_rats <- 763
mu_x <- 17.2
```

Consider question 6.6. "Rats doing mazes while on drugs". They have $N = `r N_rats` who completed the maze on the standard drug in an average of $\mu_x=`r mu_x`$ seconds. 

A random sample of 11 rats are given a new drug. Their old times $x_i$ were known from before and they complete the maze while on the new drug in time $y_i$. 

The task is to estimate the average maze time $\mu_y$ for the new drug. 

## these are your rats on drugs

```{r, echo=FALSE, fig.align='center', fig.width=3, fig.height=3}
r_rats <- mean(rats$New)/mean(rats$Standard)
rats %>% 
  ggplot(aes(x=Standard, y=New)) + geom_point() + 
  ylim(13, 20)+ 
  geom_abline(slope = r_rats, intercept=0)
rats %>% 
  select(-Rat) %>% 
  mutate(r_i = New - r_rats*Standard) -> rats_aug
s2r <- var(rats_aug$r_i)
```

The estimated ratio is $r = `r r_rats`$. The mean estimate is 
$$\hat\mu_y = 

$s^2_r = `r s2r`$. 



